Evidence for a Color Parameter Unassociated With Dust Within the Type~Ia Supernovae of the Nearby Supernova Factory

### 2017-11-29 submit/fix.tex
forging ahead

\subsection{Model}
\label{modelIImodel:sec}
Model~II is identical to Model~I except for the 
addition of an intrinsic color parameter $p$.  The distribution of this parameter over all supernovae is
assumed to be drawn from a Normal distribution with unit standard deviation.
The relative effect on each band is specified by the global parameters 
$\sigma_p\vec{\phi}$, where $\vec{\phi}$ is a unit vector.
Model~II reads as
\begin{equation}
p  \sim \mathcal{N}(0,1 ),
\end{equation}
\begin{equation}
\sigma_p  \sim \text{Cauchy}(0.1,0.1),
\end{equation}
\begin{equation}
\begin{pmatrix}
{\hat{U}}\\{\hat{B}}\\{\hat{V}}\\{\hat{R}}\\{\hat{I}}
\end{pmatrix}
=
\Delta  \vec{I} +
\begin{pmatrix}
c_{\hat{U}}+\alpha_{\hat{U}} EW_{Ca} + \beta_{\hat{U}} EW_{Si} + \eta_{\hat{U}} \lambda_{Si} + \sigma_p\phi_{\hat{U}} p\\
c_{\hat{B}}+\alpha_{\hat{B}} EW_{Ca} + \beta_{\hat{B}} EW_{Si} + \eta_{\hat{B}} \lambda_{Si}  + \sigma_p\phi_{\hat{B}} p\\
c_{\hat{V}}+\alpha_{\hat{V}} EW_{Ca} + \beta_{\hat{V}} EW_{Si} + \eta_{\hat{V}} \lambda_{Si} + \sigma_p\phi_{\hat{V}}  p\\
c_{\hat{R}}+\alpha_{\hat{R}} EW_{Ca} + \beta_{\hat{R}} EW_{Si} + \eta_{\hat{R}} \lambda_{Si} + \sigma_p\phi_{\hat{R}} p\\
c_{\hat{I}}+\alpha_{\hat{I}} EW_{Ca} + \beta_{\hat{I}} EW_{Si}+ \eta_{\hat{I}} \lambda_{Si}+ \sigma_p\phi_{\hat{I}}  p
\end{pmatrix},
\label{ewsiv2:eqn}
\end{equation}
\begin{equation}
\begin{pmatrix}
{\hat{U}}_o\\{\hat{B}}_o\\ {\hat{V}}_o\\{\hat{R}}_o\\{\hat{I}}_o\\EW_{Si, o}\\ EW_{Ca, o} \\ \lambda_{Si, o}
\end{pmatrix}
\sim \mathcal{N}
\left(
\begin{pmatrix}
{\hat{U}} +\gamma^0_{{\hat{U}}} g_0 +\gamma^1_{{\hat{U}}} g_1 \\{\hat{B}} +\gamma^0_{{\hat{B}}} g_0 +\gamma^1_{{\hat{B}}} g_1 \\
{\hat{V}}+\gamma^0_{{\hat{V}}} g_0+\gamma^1_{{\hat{V}}} g_1\\{\hat{R}}+\gamma^0_{{\hat{R}}} g_0 + \gamma^1_{{\hat{R}}} g_1\\{\hat{I}}+\gamma^0_{{\hat{I}}} g_0+\gamma^1_{{\hat{I}}} g_1\\
EW_{Si}\\ EW_{Ca} \\ \lambda_{Si}
\end{pmatrix}
,C
\right).
\label{dust2:eqn}
\end{equation}

@@

For $N$ supernovae there are $8N$ observables.  There are $3N$ spectral parameters, each of
$\Delta$, $g_0$, $g_1$ contributes $N-1$ parameters, there are $N$ parameters for $p$,  and $5 \times 7$ global coefficients.
For $N=172$ supernovae, there are 1376 observables and 1236  parameters.

\subsection{Results -- $\sigma_p \vec{\phi}$: Cofactor of the Intrinsic Color Parameter}
\label{results2:sec}
In this subsection we focus on the effect of adding the third color term  $\sigma_p \vec{\phi}$.
There is strong evidence for a third color parameter.
The 68\% credible interval for $\sigma_p$ is $0.051^{+0.005}_{-0.004}$~mag, highly incompatible with zero. 
(The credible intervals of all global parameters are given in  \S\ref{m12results:sec}.)
The posterior for $\sigma_p \vec{\phi}$ is shown in Figure~\ref{M1ev:fig}.  
The two extrema are due to the sign degeneracy for $\vec{\phi}$, each of the MCMC chains converges to one of the solutions and does
not migrate between them.
The direction of $\vec{\phi}$ is not aligned with the plane defined by
the $\vec{\gamma}$ vectors; a quadratic fraction of
 $0.088^{+0.118}_{-0.063}$ of 
 $\vec{\phi}$ lies in the $\vec{\gamma}_0$--$\vec{\gamma}_1$ plane.

@@

Model~II introduces a new parameter whose distribution comes from a Normal distribution.
It is thus interesting to compare the distributions of the $\gamma^0_{\hat{V}} g_0$ and $\gamma^1_{\hat{V}}g_1$ parameters of Model~I and Model~II,
as shown in Figure~\ref{kcomp:fig}.  (Here and in future plots, the parameter value of one supernova,
e.g.\ $\gamma^0_{\hat{V}} g_0|_0$, is subtracted
out to null our some of the correlated error between supernovae.) The apparent difference between Model~I and II is in  $\gamma^1_{\hat{V}}g_1$ term,
where it seems that a Normal distribution component in Model~I is  assigned to $p$ in Model~II.

@@

@@

Light-curve shape is an established indicator of SN~Ia diversity 
\citep{1993ApJ...413L.105P, 1996ApJ...473...88R,
1997ApJ...483..565P}
associated with color \citep{1996ApJ...473...88R,2005A&A...443..781G}.
Model~III explicitly includes light-curve shape as an additional feature used to standardize supernova colors, in contrast to
the first two models that use spectral features.
This model is of primary interest
in this article and its results are described in depth.

\subsection{Model}
\label{modelIImodel:sec}
Model~III is identical to Model~II except for the 
addition of a new term  that relates magnitudes with the SALT2 light-curve-shape parameter,
$x_1$, that is related to magnitudes with linear coefficients $\zeta$. 
The model reads as
\begin{equation}
p  \sim \mathcal{N}(0,1 ),
\end{equation}
\begin{equation}
\sigma_p  \sim \text{Cauchy}(0.1,0.1),
\end{equation}
\begin{equation}
\begin{pmatrix}
{\hat{U}}\\{\hat{B}}\\{\hat{V}}\\{\hat{R}}\\{\hat{I}}
\end{pmatrix}
=
\Delta  \vec{I} +
\begin{pmatrix}
c_{\hat{U}}+\alpha_{\hat{U}} EW_{Ca} + \beta_{\hat{U}} EW_{Si} + \eta_{\hat{U}} \lambda_{Si} +\zeta_{\hat{U}} x_1 + \sigma_p\phi_{\hat{U}} p\\
c_{\hat{B}}+\alpha_{\hat{B}} EW_{Ca} + \beta_{\hat{B}} EW_{Si} + \eta_{\hat{B}} \lambda_{Si} +\zeta_{\hat{B}} x_1 + \sigma_p\phi_{\hat{B}} p\\
c_{\hat{V}}+\alpha_{\hat{V}} EW_{Ca} + \beta_{\hat{V}} EW_{Si} + \eta_{\hat{V}} \lambda_{Si} +\zeta_{\hat{V}} x_1+ \sigma_p\phi_{\hat{V}}  p\\
c_{\hat{R}}+\alpha_{\hat{R}} EW_{Ca} + \beta_{\hat{R}} EW_{Si} + \eta_{\hat{R}} \lambda_{Si} +\zeta_{\hat{R}} x_1+ \sigma_p\phi_{\hat{R}} p\\
c_{\hat{I}}+\alpha_{\hat{I}} EW_{Ca} + \beta_{\hat{I}} EW_{Si}+ \eta_{\hat{I}} \lambda_{Si}+\zeta_{\hat{I}} x_1+ \sigma_p\phi_{\hat{I}}  p
\end{pmatrix},
\label{ewsiv3:eqn}
\end{equation}
\begin{equation}
\begin{pmatrix}
{\hat{U}}_o\\{\hat{B}}_o\\ {\hat{V}}_o\\{\hat{R}}_o\\{\hat{I}}_o\\EW_{Si, o}\\ EW_{Ca, o} \\ \lambda_{Si, o} \\ x_{1, o}
\end{pmatrix}
\sim \mathcal{N}
\left(
\begin{pmatrix}
{\hat{U}} +\gamma^0_{{\hat{U}}} g_0 +\gamma^1_{{\hat{U}}} g_1 \\{\hat{B}} +\gamma^0_{{\hat{B}}} g_0 +\gamma^1_{{\hat{B}}} g_1 \\
{\hat{V}}+\gamma^0_{{\hat{V}}} g_0+\gamma^1_{{\hat{V}}} g_1\\{\hat{R}}+\gamma^0_{{\hat{R}}} g_0 + \gamma^1_{{\hat{R}}} g_1\\{\hat{I}}+\gamma^0_{{\hat{I}}} g_0+\gamma^1_{{\hat{I}}} g_1\\
EW_{Si}\\ EW_{Ca} \\ \lambda_{Si} \\ x_1
\end{pmatrix}
,C
\right).
\label{dust3:eqn}
\end{equation}
The changes from Model~II are the addition of the $\vec{\zeta} x_1$ terms in Eq.~\ref{ewsiv3:eqn}, and
the addition of the $x_1$ observation, parameter, and its implicit inclusion in the covariance matrix $C$, in Eq.~\ref{dust3:eqn}.

For $N$ supernovae there are $9N$ observables.  There are $4N$ spectral and light-curve shape parameters, each of
$\Delta$, $g_0$, $g_1$ contributes $N-1$ parameters, there are $N$ parameters for $p$,  and $5 \times 8$ global coefficients.
For $N=172$ supernovae, there are 1548 observables and 1413  parameters.
}

@@

In the analysis of this and other sections of this article,
the posterior of the model parameters is evaluated using Hamiltonian Monte Carlo with a No-U-Turn
Sampler as implemented in
STAN \citep{stan}.  We run eight chains, each with 5000 iterations of which
half are used for warmup.
Tests have revealed no evidence for the non-convergence of the MCMC run.
STAN provides output statistics to assess
the convergence of the output Markov chains.
The 
potential scale reduction statistic, $\hat{R}$
(in this paragraph not to be confused with the synthetic $R$ magnitude)
\citep{Gelman92}, measures the convergence of the target distribution
in iterative simulations 
by using multiple independent sequences to estimate how much that distribution would sharpen if the simulations were run longer.
$N_{\mathit{eff}}$ is an estimate of the number of independent draws. The output for our data and model gives $\hat{R} \sim 1.0$ for all parameters, meaning there is no evidence for non-convergence.  The
output also gives  $N_{\mathit{eff}} \gg 100$ for all parameters, indicating that all are densely sampled.

The algorithm is extensively tested with simulated data with known input signals and measurement uncertainties similar to those of our dataset.
Stacks of the posteriors from 100 simulated datasets  exhibit no biases that are significant relative to the statistical uncertainties of a single posterior.

The posterior that will be shown in  \S\ref{results3global:sec} exhibits a 
confined peak, exhibiting no evidence that the MCMC has not failed to converge.

The model provides a fair representation of the data, as seen in the difference between the data and the model prediction.
Figure~\ref{residual:fig} has plots of the observed  versus
model-expectation colors (including  supernova-specific spectral features, light-curve shape $\gamma g$, and $\phi p$ terms) for four selected colors,
and the histogram of their differences.  
There are no extreme residuals 
from catastrophically poor color predictions.  There are no apparent trends between residuals and observed
color.   The residuals of the data away from the model have a compact Gaussian-like appearance.

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=4in]{../output_fix3_x1/residual.pdf} 
            \caption{Observed  versus
            Model~III-expectation colors (including  supernova-specific spectral features, $\gamma g$, and $\phi p$  terms)
for four selected colors. 
            Plotted on the right for each color
is the histogram of the difference between the expected and
observed
 colors. 
            \label{residual:fig}}
\end{figure}

@@

\subsubsection{Global Parameter Posterior Results}
\label{results3global:sec}
Results from our analysis are shown through contours of the posterior surface for pairs of global parameters grouped by filter, 
in Figures~\ref{global1:fig} -- \ref{global5:fig}.
The confidence regions are localized and unimodal, except for the combination $\sigma_p \vec{\phi}$ that
has the sign degeneracy already noted in \S\ref{results2:sec} and plotted in Figure~\ref{M1ev:fig}.
Within the finite number of generated links, the MCMC chains do not migrate to the parity-degenerate solution of $\vec{\gamma}$ and $g$.
Each chain converges to one of the  $\vec{\phi}$--$p$ degenerate solutions, away from which it does not migrate.

@@

For each of the five filters, the 68\%  equal-tailed credible intervals for the global parameters $\vec{\alpha}$, $\vec{\beta}$, $\vec{\eta}$, $\vec{\zeta}$,
$\vec{\gamma}^0$, $\vec{\gamma}^1$, and $\sigma_p\vec{\phi}$
are given in Table~\ref{global2:tab}.
To break the sign degeneracy, we enforce that $\phi_{\hat{I}}$
be positive such that  $\vec{\phi} = \text{sign}(\phi_{\hat{I}}) \vec{\phi}$;
we use the $\hat{I}$-band because
Figure~\ref{global5:fig} shows that it has
the largest separation between the degenerate solutions of $\sigma_p \phi_X$.
The effect of spectral parameters on color (as opposed to magnitude)
is shown in the rows of $\alpha_X/\alpha_{\hat{V}}-1$,  $\beta_X/\beta_{\hat{V}}-1$, and  $\eta_X/\eta_{\hat{V}}-1$.
The freedom to scale each of the terms of $\vec{\gamma}g$ is nulled out in the statistic
 $\gamma^i_X/\gamma^i_{\hat{V}}-1$.
Note the analogy between these statistics and the common use of
$R_V^{-1}=A_B/A_V-1$ as an observational descriptor of dust properties.

>>>>

\subsection{Model}
\label{modelIImodel:sec}
Model~II is identical to Model~I except for the 
addition of an intrinsic color parameter $p$.  The distribution of this parameter over all supernovae is
assumed to be drawn from a Normal distribution with unit standard deviation; leaving the distribution unconstrained,
as was done for $g_0$ and $g_1$, yields many local maxima in the posterior and makes MCMC convergence difficult.
The relative effect on each band is specified by the global parameters 
$\sigma_p\vec{\phi}$, where $\vec{\phi}$ is a unit vector.
Model~II reads as
\begin{equation}
p  \sim \mathcal{N}(0,1 ),
\end{equation}
\begin{equation}
\sigma_p  \sim \text{Cauchy}(0.1,0.1),
\end{equation}
\begin{equation}
\begin{pmatrix}
{\hat{U}}\\{\hat{B}}\\{\hat{V}}\\{\hat{R}}\\{\hat{I}}
\end{pmatrix}
=
\Delta  \vec{I} +
\begin{pmatrix}
c_{\hat{U}}+\alpha_{\hat{U}} EW_{Ca} + \beta_{\hat{U}} EW_{Si} + \eta_{\hat{U}} \lambda_{Si} + \sigma_p\phi_{\hat{U}} p\\
c_{\hat{B}}+\alpha_{\hat{B}} EW_{Ca} + \beta_{\hat{B}} EW_{Si} + \eta_{\hat{B}} \lambda_{Si}  + \sigma_p\phi_{\hat{B}} p\\
c_{\hat{V}}+\alpha_{\hat{V}} EW_{Ca} + \beta_{\hat{V}} EW_{Si} + \eta_{\hat{V}} \lambda_{Si} + \sigma_p\phi_{\hat{V}}  p\\
c_{\hat{R}}+\alpha_{\hat{R}} EW_{Ca} + \beta_{\hat{R}} EW_{Si} + \eta_{\hat{R}} \lambda_{Si} + \sigma_p\phi_{\hat{R}} p\\
c_{\hat{I}}+\alpha_{\hat{I}} EW_{Ca} + \beta_{\hat{I}} EW_{Si}+ \eta_{\hat{I}} \lambda_{Si}+ \sigma_p\phi_{\hat{I}}  p
\end{pmatrix},
\label{ewsiv2:eqn}
\end{equation}
\begin{equation}
\begin{pmatrix}
{\hat{U}}_o\\{\hat{B}}_o\\ {\hat{V}}_o\\{\hat{R}}_o\\{\hat{I}}_o\\EW_{Si, o}\\ EW_{Ca, o} \\ \lambda_{Si, o}
\end{pmatrix}
\sim \mathcal{N}
\left(
\begin{pmatrix}
{\hat{U}} +\gamma^0_{{\hat{U}}} g_0 +\gamma^1_{{\hat{U}}} g_1 \\{\hat{B}} +\gamma^0_{{\hat{B}}} g_0 +\gamma^1_{{\hat{B}}} g_1 \\
{\hat{V}}+\gamma^0_{{\hat{V}}} g_0+\gamma^1_{{\hat{V}}} g_1\\{\hat{R}}+\gamma^0_{{\hat{R}}} g_0 + \gamma^1_{{\hat{R}}} g_1\\{\hat{I}}+\gamma^0_{{\hat{I}}} g_0+\gamma^1_{{\hat{I}}} g_1\\
EW_{Si}\\ EW_{Ca} \\ \lambda_{Si}
\end{pmatrix}
,C
\right).
\label{dust2:eqn}
\end{equation}

@@

For $N$ supernovae there are $8N$ observables.  There are $3N$ spectral parameters, each of
$\Delta$, $g_0$, $g_1$ contributes $N-1$ parameters while $p$  contributes $N$ parameters,  and  there are $5 \times 7$ global coefficients.
For $N=172$ supernovae, there are 1376 observables and 1236  parameters.

\subsection{Results -- $\sigma_p \vec{\phi}$: Cofactor of the Intrinsic Color Parameter}
\label{results2:sec}
In this subsection we focus on the effect of adding the third color term  $\sigma_p \vec{\phi}$.
There is strong evidence for a third color parameter.
The 68\% credible interval for $\sigma_p$ is $0.051^{+0.005}_{-0.004}$~mag, highly incompatible with zero. 
The posterior for $\sigma_p \vec{\phi}$ is shown in Figure~\ref{M1ev:fig}.  
The two maxima are due to the sign degeneracy for $\vec{\phi}$, each of the MCMC chains converges to one of the solutions and does
not migrate between them.
The direction of $\vec{\phi}$ is not aligned with the plane defined by
the $\vec{\gamma}$ vectors; only a quadratic fraction of
 $0.088^{+0.118}_{-0.063}$ of 
 $\vec{\phi}$ lies in the $\vec{\gamma}_0$--$\vec{\gamma}_1$ plane.

@@

Model~II introduces a new parameter  that can absorb color varability
that was restricted to $g_0$ and $g_1$ in Model~I, and whose distribution comes from a Normal distribution.
The change in the distributions of the $\gamma^0_{\hat{V}} g_0$ and $\gamma^1_{\hat{V}}g_1$ parameters from
 Model~I to Model~II
is apparent in Figure~\ref{kcomp:fig}.  (Here and in future plots, the parameter value of one supernova,
e.g.\ $\gamma^0_{\hat{V}} g_0|_0$, is subtracted
out to null out correlated error between supernovae.) The  change from Model~I to II is pronounced in  $\gamma^1_{\hat{V}}g_1$,
where it seems that a Normal distribution component in Model~I disappears in Model~II, presumably having been  assigned to $p$.

@@

The credible intervals of all Model~II global parameters are given in  \S\ref{m12results:sec}.

@@

Light-curve shape is an established indicator of SN~Ia diversity 
\citep{1993ApJ...413L.105P, 1996ApJ...473...88R,
1997ApJ...483..565P}
with its own color correlations \citep{2005A&A...443..781G, 2007ApJ...659..122J}.
Model~III explicitly includes light-curve shape as a feature used to standardize supernova colors, in addition to the
spectral features considered in the previous two models.
We will show that light-curve shape does have a non-trivial relationship with color that is not already encoded
in the spectral features.  Model~III is thus the focus of
this article and its results are  described here in depth.

\subsection{Model}
\label{modelIImodel:sec}
Model~III is identical to Model~II except for the 
addition of a new term  that relates magnitudes with the SALT2 light-curve-shape parameter,
$x_1$, through linear coefficients $\zeta$. 
The model reads as
\begin{equation}
p  \sim \mathcal{N}(0,1 ),
\end{equation}
\begin{equation}
\sigma_p  \sim \text{Cauchy}(0.1,0.1),
\end{equation}
\begin{equation}
\begin{pmatrix}
{\hat{U}}\\{\hat{B}}\\{\hat{V}}\\{\hat{R}}\\{\hat{I}}
\end{pmatrix}
=
\Delta  \vec{I} +
\begin{pmatrix}
c_{\hat{U}}+\alpha_{\hat{U}} EW_{Ca} + \beta_{\hat{U}} EW_{Si} + \eta_{\hat{U}} \lambda_{Si} +\zeta_{\hat{U}} x_1 + \sigma_p\phi_{\hat{U}} p\\
c_{\hat{B}}+\alpha_{\hat{B}} EW_{Ca} + \beta_{\hat{B}} EW_{Si} + \eta_{\hat{B}} \lambda_{Si} +\zeta_{\hat{B}} x_1 + \sigma_p\phi_{\hat{B}} p\\
c_{\hat{V}}+\alpha_{\hat{V}} EW_{Ca} + \beta_{\hat{V}} EW_{Si} + \eta_{\hat{V}} \lambda_{Si} +\zeta_{\hat{V}} x_1+ \sigma_p\phi_{\hat{V}}  p\\
c_{\hat{R}}+\alpha_{\hat{R}} EW_{Ca} + \beta_{\hat{R}} EW_{Si} + \eta_{\hat{R}} \lambda_{Si} +\zeta_{\hat{R}} x_1+ \sigma_p\phi_{\hat{R}} p\\
c_{\hat{I}}+\alpha_{\hat{I}} EW_{Ca} + \beta_{\hat{I}} EW_{Si}+ \eta_{\hat{I}} \lambda_{Si}+\zeta_{\hat{I}} x_1+ \sigma_p\phi_{\hat{I}}  p
\end{pmatrix},
\label{ewsiv3:eqn}
\end{equation}
\begin{equation}
\begin{pmatrix}
{\hat{U}}_o\\{\hat{B}}_o\\ {\hat{V}}_o\\{\hat{R}}_o\\{\hat{I}}_o\\EW_{Si, o}\\ EW_{Ca, o} \\ \lambda_{Si, o} \\ x_{1, o}
\end{pmatrix}
\sim \mathcal{N}
\left(
\begin{pmatrix}
{\hat{U}} +\gamma^0_{{\hat{U}}} g_0 +\gamma^1_{{\hat{U}}} g_1 \\{\hat{B}} +\gamma^0_{{\hat{B}}} g_0 +\gamma^1_{{\hat{B}}} g_1 \\
{\hat{V}}+\gamma^0_{{\hat{V}}} g_0+\gamma^1_{{\hat{V}}} g_1\\{\hat{R}}+\gamma^0_{{\hat{R}}} g_0 + \gamma^1_{{\hat{R}}} g_1\\{\hat{I}}+\gamma^0_{{\hat{I}}} g_0+\gamma^1_{{\hat{I}}} g_1\\
EW_{Si}\\ EW_{Ca} \\ \lambda_{Si} \\ x_1
\end{pmatrix}
,C
\right).
\label{dust3:eqn}
\end{equation}
The changes from Model~II are the addition of the $\vec{\zeta} x_1$ terms in Eq.~\ref{ewsiv3:eqn}, and
the addition of the shape measurement and uncertainty (implicit in the covariance matrix $C$) and $x_1$   in Eq.~\ref{dust3:eqn}.

For $N$ supernovae there are $9N$ observables.  There are $4N$ spectral and light-curve shape parameters, each of
$\Delta$, $g_0$, $g_1$ contributes $N-1$ parameters while $p$  contributes $N$ parameters,  and there are $5 \times 8$ global coefficients.
For $N=172$ supernovae, this makes 1548 observables and 1413  parameters.

@@

In the analysis of this and previous sections of this article,
the posterior of the model parameters is evaluated using Hamiltonian Monte Carlo with a No-U-Turn
Sampler as implemented in
STAN \citep{JSSv076i01}.  We run eight chains, each with 5000 iterations of which
half are used for warmup.
STAN provides output statistics to assess
the convergence of the output Markov chains.
The 
potential scale reduction statistic, $\hat{R}$
\citep[][in this paragraph not to be confused with the synthetic $\hat{R}$-band magnitude]{Gelman92}, measures the convergence of the target distribution
in iterative simulations 
by using multiple independent sequences to estimate how much that distribution would sharpen if the simulations were run longer.
$N_{\mathit{eff}}$ is an estimate of the number of independent draws. The STAN output gives $\hat{R} \sim 1.0$ for all parameters, meaning there is no evidence for non-convergence.  The
output also gives  $N_{\mathit{eff}} \gg 100$ for all parameters, indicating that they are densely sampled.

The algorithm is run with input simulated data with known  signals and measurement uncertainties similar to those of our dataset.
Stacks of the posteriors from 100 simulated datasets  exhibit no biases that are significant relative to the statistical uncertainties of a single posterior.

As will be seen in  \S\ref{results3global:sec}, the final posterior has a distinct bounded  
peak showing no indication that the MCMC has failed to converge.

The model provides a fair representation of the data as seen in the residual differences between the data and the model prediction.
Figure~\ref{residual:fig} shows plots of the observed  versus
model-expectation colors (differences between band magnitudes in the mean of the Normal distribution of Eq.~\ref{dust3:eqn}) for four selected colors,
and the histogram of the residuals.  The residuals have compact Gaussian-like appearance with
no catastrophically poor color predictions.  There are no apparent trends between residuals and observed
color.

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=4in]{../output_fix3_x1/residual.pdf} 
            \caption{Observed  versus
            Model~III-expectation colors (including  contributions related to spectral features, light-curve shape, $\gamma g$, and $\phi p$  terms)
for four selected colors. 
            Plotted on the right for each color
is the histogram of the difference between the expected and
observed
 colors. 
            \label{residual:fig}}
\end{figure}

@@

\subsubsection{Global Parameter Posterior Results}
\label{results3global:sec}
Results from our analysis are shown in Figures~\ref{global1:fig} -- \ref{global5:fig}
as contours of the posterior surface for pairs of global parameters grouped by filter.
The confidence regions are localized and unimodal, except for the combination $\sigma_p \vec{\phi}$ that
has the sign degeneracy explained in \S\ref{results2:sec} and plotted in Figure~\ref{M1ev:fig}.
Within the finite number of generated links, the MCMC chains do not migrate between the parity-degenerate $\vec{\gamma}$--$g$ solutions.
Each chain converges to one of the  $\vec{\phi}$--$p$ degenerate solutions, away from which it does not migrate.

@@

For each of the five filters, the 68\%  equal-tailed credible intervals for the global parameters $\vec{\alpha}$, $\vec{\beta}$, $\vec{\eta}$, $\vec{\zeta}$,
$\vec{\gamma}^0$, $\vec{\gamma}^1$, and $\sigma_p\vec{\phi}$
are given in Table~\ref{global2:tab}.
In constructing the credible interval of $\sigma_p\vec{\phi}$ 
we set $\vec{\phi} = \text{sign}(\phi_{\hat{I}}) \vec{\phi}$ to break its parity degeneracy;
Figure~\ref{global5:fig} shows there the degenerate solutions are relatively well separated 
in $\sigma_p \phi_{\hat{I}}$.
The effect of spectral parameters on color (as opposed to magnitude)
is shown in the rows of $\alpha_X/\alpha_{\hat{V}}-1$,  $\beta_X/\beta_{\hat{V}}-1$, and  $\eta_X/\eta_{\hat{V}}-1$.
The normalization freedom of $\vec{\gamma}$ is nulled out in the statistic
 $\gamma_X/\gamma_{\hat{V}}-1$.
Note the analogy between these color statistics and the common use of
$R_V^{-1}=A_B/A_V-1$ as an observational descriptor of dust properties.

\begin{table}
\centering
\begin{tabular}{|c|c|c|c|c|c|}
\hline
Parameters& $X={\hat{U}}$ &${\hat{B}}$&${\hat{V}}$&${\hat{R}}$&${\hat{I}}$\\ \hline
$\alpha_X$
& $0.0059^{+0.0007}_{-0.0008}$
& $0.0033^{+0.0006}_{-0.0007}$
& $0.0029^{+0.0006}_{-0.0006}$
& $0.0026^{+0.0005}_{-0.0005}$
& $0.0041^{+0.0004}_{-0.0005}$
\\
${\alpha_X/\alpha_{\hat{V}}-1}$
& $   1.0^{+   0.2}_{  -0.2}$
& $   0.1^{+   0.1}_{  -0.1}$
&\dots
& $  -0.1^{+   0.0}_{  -0.0}$
& $   0.4^{+   0.2}_{  -0.1}$
\\
$\beta_X$
& $ 0.032^{+ 0.006}_{-0.006}$
& $ 0.021^{+ 0.005}_{-0.005}$
& $ 0.018^{+ 0.005}_{-0.005}$
& $ 0.016^{+ 0.004}_{-0.004}$
& $ 0.008^{+ 0.004}_{-0.004}$
\\
${\beta_X/\beta_{\hat{V}}-1}$
& $  0.79^{+  0.29}_{ -0.19}$
& $  0.16^{+  0.11}_{ -0.08}$
&\dots
& $ -0.12^{+  0.04}_{ -0.04}$
& $ -0.58^{+  0.11}_{ -0.17}$
\\
$\eta_X$
& $0.0022^{+0.0011}_{-0.0011}$
& $0.0024^{+0.0009}_{-0.0010}$
& $0.0027^{+0.0008}_{-0.0008}$
& $0.0025^{+0.0007}_{-0.0007}$
& $0.0018^{+0.0006}_{-0.0006}$
\\
${\eta_X/\eta_{\hat{V}}-1}$
& $ -0.19^{+  0.17}_{ -0.28}$
& $ -0.12^{+  0.09}_{ -0.15}$
&\dots
& $ -0.06^{+  0.06}_{ -0.04}$
& $ -0.33^{+  0.10}_{ -0.10}$
\\
$\zeta_X$
& $ -0.06^{+  0.05}_{ -0.04}$
& $ -0.09^{+  0.04}_{ -0.04}$
& $ -0.12^{+  0.04}_{ -0.04}$
& $ -0.09^{+  0.03}_{ -0.03}$
& $ -0.14^{+  0.03}_{ -0.03}$
\\
${\zeta_X/\zeta_{\hat{V}}-1}$
& $ -0.45^{+  0.20}_{ -0.33}$
& $ -0.26^{+  0.11}_{ -0.18}$
&\dots
& $ -0.20^{+  0.05}_{ -0.07}$
& $  0.22^{+  0.23}_{ -0.13}$
\\
$\gamma^0_X$
& $ 68.29^{+  2.97}_{ -2.92}$
& $ 54.90^{+  2.49}_{ -2.56}$
& $ 40.89^{+  2.38}_{ -2.47}$
& $ 30.95^{+  2.13}_{ -2.25}$
& $ 22.34^{+  2.05}_{ -2.16}$
\\
${\gamma^0_X/\gamma^0_{\hat{V}}-1}$
& $  0.67^{+  0.05}_{ -0.05}$
& $  0.34^{+  0.03}_{ -0.03}$
&\dots
& $ -0.24^{+  0.01}_{ -0.01}$
& $ -0.45^{+  0.03}_{ -0.03}$
\\
$\gamma^1_X$
& $-10.15^{+  4.45}_{ -4.49}$
& $ -6.81^{+  3.88}_{ -3.82}$
& $-15.06^{+  3.42}_{ -3.39}$
& $-14.88^{+  3.01}_{ -3.00}$
& $-15.91^{+  2.90}_{ -2.86}$
\\
${\gamma^1_X/\gamma^1_{\hat{V}}-1}$
& $ -0.32^{+  0.16}_{ -0.21}$
& $ -0.55^{+  0.13}_{ -0.20}$
&\dots
& $ -0.01^{+  0.05}_{ -0.04}$
& $  0.06^{+  0.14}_{ -0.10}$
\\
$\sigma_p \phi_X$
& $ 0.024^{+ 0.009}_{-0.009}$
& $-0.018^{+ 0.007}_{-0.007}$
& $-0.027^{+ 0.006}_{-0.006}$
& $-0.010^{+ 0.006}_{-0.006}$
& $ 0.022^{+ 0.006}_{-0.006}$
\\
\hline
\end{tabular}
\caption{68\% credible intervals for the global fit parameters of Model~III in \S\ref{modelIII:sec}.\label{global3:tab}}
\end{table}

\subsubsection{Results for the Observables Cofactors $\vec{\alpha}$, $\vec{\beta}$, $\vec{\eta}$,  $\vec{\zeta}$}

The cofactors  $\vec{\alpha}$, $\vec{\beta}$, and $\vec{\eta}$ have  significant non-zero values,  meaning that
$EW_{Ca}$, $EW_{Si}$, and $\lambda_{Si}$ are indicators of broadband
colors at peak.  The significance of their influence varies between  colors; in $\hat{B}-\hat{V}$, roughly the color
where previous studies have focused, only $\vec{\beta}$ has a strong $>2\sigma$ correlation.  A positive
dependence (consistent with the positive sign
of $\beta_{\hat{B}}/\beta_{\hat{V}}-1$)
of $B-V$ on the pseudo-equivalent width of Si~II~$\lambda$4130 has previously been reported by
\citet{2013ApJ...773...53F}.
The insignificance of  our $\alpha_{\hat{B}}/\alpha_{\hat{V}}-1$ is consistent with the
lack of correlation between the  pseudo-equivalent width of Ca~II~H\&K measurements and
intrinsic color reported by \citet{2011ApJ...742...89F}.

The story on the Si~II~$\lambda$6355 is not as clear.
 \citet{2011ApJ...742...89F, 2012ApJ...748..127F} find that
higher-velocity SNe~Ia tend to be redder than those with lower velocity. \citet{2012AJ....143..126B} find a
weaker relation than that reported by \citet{2011ApJ...742...89F},
a difference attributed to a better treatment of uncertainties and host-galaxy dust. In contrast,
\citet{2013ApJ...773...53F} find no significant relation.
Our result for $\eta_{\hat{B}}/\eta_{\hat{V}}-1$ also gives no significant relation.

The non-zero values for $\vec{\zeta}$ demonstrate that light-curve shape $x_1$ is also an indicator
of broadband colors, in the sense that SNe~Ia with broader light curves are bluer in all colors.
This is as expected based on the results of \citet{2007A&A...466...11G} and MLCS2k2
\citep{2007ApJ...659..122J}.
}

The signal in  $\vec{\alpha}$ and $\vec{\beta}$ cannot
entirely
be attributed to the equivalent widths themselves.
The range of Si~II~$\lambda$4130 equivalent widths is $\pm 20$~\AA\ whereas the width of the ${\hat{B}}$-band is 851~\AA, so that its direct affect on magnitude
is
$2.5 \log{(20/850)} \sim 0.03$ mag.  
The implied span in ${\hat{B}}$ magnitude based on $\beta_{\hat{B}}$ is 0.54~mag.  Therefore $\beta_{\hat{B}}$ cannot wholly be attributed to the flux deficit
from the line itself.
The Ca~II H\&K equivalent widths have range $\pm 50$~\AA, while the width of the ${\hat{U}}$ band is
701~\AA, so that its direct affect on magnitude
is
$2.5 \log{(50/701)} \sim 0.08$ mags.   The implied span in ${\hat{U}}$ magnitude is  0.21~mag, 
which would imply that  $\alpha_{\hat{U}}$ cannot be completely due to the 
presence of the line itself.  
Supernova flux has a large gradient in our $\hat{U}$ and there is a variation in the effective, flux-weighted,  $\hat{U}$ bandwidth. 
Nevertheless, the  correlation between spectral features with magnitudes in bands with no wavelength overlap is a non-trivial
signal detected in the analysis.
